### 2023-08-10 MSCA_31006_Time_Series_Final_Project.Rmd
Add files via upload

---
title: 'MSCA 31006: Final Project - AQI Forecasting'
author: "Aashish Singh"
date: "8/10/2023"
output: html_document
---

```{r, tidy=TRUE}
#install.packages(c("forecast", "expsmooth", "seasonal")) 
library(TTR)
library(forecast) 
library(tseries) 
library(expsmooth) 
library(fpp2)
library(seasonal)
library(MASS)
library(TSA)
library(forecast)
library(ggplot2)
library(tseries)
library(imputeTS)
library(vars)
library(timetk)
library(Metrics)
library(lmtest)
library(lubridate)
library(xts)
library(dplyr)
```

#### Step 1: Load data and plot the time series

```{r, tidy=TRUE}
path <- "/Users/aashishsingh/Desktop/Time Series - Final Project/"
fresno_pm2_5 <- read.csv(paste0(path, "fresno_pm2_5_pollutant_data.csv"), 
                         na.strings=c("", "NA"))
fresno_ozone <- read.csv(paste0(path, "fresno_ozone_pollutant_data.csv"), 
                         na.strings=c("", "NA"))

# Convert data into ts objects
fresno_pm2_5_ts <- ts(fresno_pm2_5$PM2.5.AQI.Value, start = c(1999,1,1), frequency = 365.25)
fresno_ozone_ts <- ts(fresno_ozone$Ozone.AQI.Value, start = c(1999,1,1), frequency = 365.25)

plot(fresno_pm2_5_ts, main="Fresno PM 2.5 AQI")
plot(fresno_ozone_ts, main="Fresno Ozone AQI")
```

#### Step 2: Check if data is stationary (KPSS/ADF Test)

```{r, tidy=TRUE}

kpss.test(fresno_pm2_5_ts)
# The reported p-value is 0.01, which is smaller than 0.05, and would suggest 
# that we reject the null hypothesis of level stationarity and conclude that 
# there is evidence that the data is non-stationary

adf.test(fresno_pm2_5_ts)
# Similarly the Augumented Dickey-Fuller test results in p-value of 0.01 (<0.05) 
# where we do reject the null hypothesis that the time series is non-stationary
# and thus data is stationary.

# These are opposite results so we use ACF PACF plots to see if the series is stationary
acf(fresno_pm2_5_ts, main = "ACF of Original Time Series")
pacf(fresno_pm2_5_ts, main = "PACF of Original Time Series")

```

#### Step 3: Extract weekly values for time series and plot the series

```{r, tidy=TRUE}
# Create weekly data
fresno_pm2_5_df <- as.data.frame(fresno_pm2_5_ts)
fresno_pm2_5_df$Date <- mdy(fresno_pm2_5$Date)
colnames(fresno_pm2_5_df) <- c("PM2.5.AQI.Value", "Date")

fresno_pm2_5_xts <- as.xts(fresno_pm2_5_df, order.by=fresno_pm2_5_df$Date)
fresno_pm2_5_xts <- fresno_pm2_5_xts[,-2]
fresno_pm2_5_weekly <- apply.weekly(fresno_pm2_5_xts, mean)

fresno_pm2_5_weekly_ts <- ts(fresno_pm2_5_weekly["19990103/20230805"], start = c(1999,1),frequency = 52.18)
plot(fresno_pm2_5_weekly_ts)

kpss.test(fresno_pm2_5_weekly_ts)

adf.test(fresno_pm2_5_weekly_ts)

shapiro.test(fresno_pm2_5_weekly_ts)

lambda <- BoxCox.lambda(fresno_pm2_5_weekly_ts)

ts.plot(BoxCox(fresno_pm2_5_weekly_ts, lambda), main="Box-Cox Transformed PM 2.5 AQI")

fresno_pm2_5_weekly_ts_transformed <- BoxCox(fresno_pm2_5_weekly_ts, lambda)

kpss.test(fresno_pm2_5_weekly_ts_transformed)

adf.test(fresno_pm2_5_weekly_ts_transformed)

acf(fresno_pm2_5_weekly_ts_transformed, main = "ACF of Box-Cox Transformed Time Series")
pacf(fresno_pm2_5_weekly_ts_transformed, main = "PACF of Box-Cox Transformed Time Series")

```

#### Step 4: Extract monthly values for time series and plot the series

```{r, tidy=TRUE}
# Create monthly data
fresno_pm2_5_monthly <- apply.monthly(fresno_pm2_5_xts, mean)

fresno_pm2_5_monthly_ts <- ts(fresno_pm2_5_monthly["19990130/20230805"], start = c(1999,1), frequency = 12)
plot(fresno_pm2_5_monthly_ts)

kpss.test(fresno_pm2_5_monthly_ts)

adf.test(fresno_pm2_5_monthly_ts)

acf(fresno_pm2_5_monthly_ts, main = "ACF of Monthly Time Series")
pacf(fresno_pm2_5_monthly_ts, main = "PACF of Monthly Time Series")
```